### 2017-04-26 src/hardness.tex
going to sleep

In this section we show that the problem in its general form cannot be
approximated at all assuming $P \neq NP$.
We also show that the problem remains NP-hard even when restricted to DAGS.

Consider the problem in its general form. 
Given an undirected graph $G$