### 2017-04-26 src/hardness.tex

In this section we show that the problem in its general form cannot be
approximated at all assuming $P \neq NP$.
\todo{Is it true ?}
We also show that the problem remains NP-hard even when restricted to DAGS.

Consider the problem in its general form. 
Given an undirected graph $G$
\todo{finish}

We now show that the problem restricted to DAG is still NP-hard.
We present a polynomial reduction from partition.
Let $\{a_1, \ldots, a_n\}$ be an instance of the partition problem, 
consider the instance of the VPN problem 
$D_v = (V_v, A_v), D_p = ({v}, \emptyset)$, 
where:
\begin{align}
V_v = \{s, t\} \cup \{v_1^+, \ldots, v_n^+\} \cup \{v_1^-, \ldots, v_n^-\}
\\
A_v = \{\{v_i^+, v_i^-\} \times \{v_{i + 1}^+, v_{i + 1}^-\}\} 
\cup \{(s, v_1^+), (s, v_1^-), (v_n^+, t), (v_n^-, t)\}
&&
\forall i \in [n-1]
\\
p_d(v_i^+) = a_i && \forall i \in [n]
\\
p_d(v_i^-) = 0 && \forall i \in [n]
\\
c(v_i^+, v) = 0 && \forall i \in [n]
\\
c(v_i^-, v) = a_i && \forall i \in [n] 
\\
p_c(v) = \inv{2}\sum_{i = 1}^n a_i
\end{align}
Note that the bandwidth function does not plays a role here because the physical graph
is just a single vertex.

The following figure illustrate the reduction.
\begin{figure}[ht]
\centering
\input{fig-reduction}
\end{figure}