### 2017-04-26 src/main.tex

\nonstopmode
\documentclass[draft]{article}

\input{libs}
\input{def}
\input{tikz}

\input{title}
\input{authors}

\begin{document}
\maketitle

\input{outline}

\bibliographystyle{plain}
\bibliography{main}

\end{document}