### 2017-04-26 src/latency.tex
done for today

\todo{Are we going to talk about the problem also with latency constraint ?}