### 2017-04-27 src/fig-dp.tex
latency

\begin{tikzpicture}[every node/.style={default node}, very thick, -latex]

\node[clear node] at(-3, 0) {$\dots$};

\node(s) at(-2,0) {$a_v$};

\node(c1) at(1,1) {$c_1$};
\node[clear node] at (2,0) {$\vdots$};
\node(c2) at(1,-1) {$c_d$};

\node(t) at(5, 0) {$t'_v$};

\node[clear node] at(6, 0) {$\dots$};

\draw (s) to[] (c1);
\draw (s) to[]
node[clear node, pos=0.5](c1n){}
(c2);

\begin{scope}[path]
\draw[decorate] (c1.north) to[out=90, in=90] (t.north east);
\draw[decorate] (c1) to[out=45, in=135] (t);
\draw[decorate] (c1.south east) to[out=0, in=135] (t.south west);

\draw[decorate] (c2.north east) to[out=0, in=-135] (t.north west);
\draw[decorate] (c2) to[out=-45, in=-135] (t);
\draw[decorate] (c2.south) to[out=-90,in=-90] (t.south east);
\end{scope}

\draw[group]
(c2.west)		to[out=-90,in=-90, looseness=1.8] node[clear node, pos=0.9](j){}
(t.east)		to[out=90,in=0]
(t.north)		to[out=180,in=45]
(c2.north west)	to[out=-135,in=90]
(c2.west)
;

\node[clear node] at (-3,-5) {$\dots$};

\node(sp) at(-2, -5) {$s'_p$};
\node(tp) at(5, -5) {$t'_p$};

\node[clear node] at (6,-5) {$\dots$};

\begin{scope}[path]
\draw[decorate]	(sp.north east) to[bend left] 
node[pos=0.5, clear node](p1){}
(tp.north west)
;
\draw[decorate] (sp) to[] 
node[pos=0.3, clear node](p2){}
(tp);
\draw[decorate]	(sp.south east) to[bend right] (tp.south west);
\end{scope}

\draw[group, ->] (j) to[bend left] (tp);

\begin{scope}[blue, -Rays, every to/.style={bend right}]
\draw (c1n) to[] (p1);
\draw (c1n) to[] (p2);
\end{scope}
\end{tikzpicture}