### 2017-04-27 src/latency.tex

\todo{Are we going to talk about the problem also with latency constraint ?}
We now extend the problem by adding a global latency constraint.
Formally, we are given a latency function $l : A_v \times A_p \to \R$,
and a latency constraint $L$.
The goal is to find a min-cost embedding that also respect the latency
constraint.
Observe that given two physical vertices $s_p$, $t_p$ and a virtual edge $e_v$
a $s_pt_p$-path that minimize the latency can be efficiently found by using any
standard shortest path algorithm where the "distance" function is replaced by
the appropriate latency.
Observe that this can also be done if we restrict ourselves to use only physical
edges with a lower bound on their bandwidth.
We now show that if all values are integral and polynomial bounded, we can
efficiently find a feasible embedding with a cost bounded by $C$ that minimize
the latency.

Let $O(t'_v, t'_p, C)$ be an embedding of some virtual $s_vt'_v$-path 
into a physical $s_pt'_p$-path that has cost not more than $C$ and that
minimizes the latency.