### 2017-04-28 src/problem.tex

We are given a virtual directed graph $D_v = (V_v, A_v)$ 
with source and destination vertices $s_v, t_v$ respectively, 
a physical directed graph $D_p = (V_p, A_p)$
with source and destination vertices $s_p, t_p$ respectively,
demand functions $p_d : V_v \to \R$, $b_d : E_v \to \R$,
capacity functions $p_c : V_p \to \R$, $b_c : E_p \to \R$,
and a cost function $c: V_v \times V_p \to \R$.

Given a virtual $s_vt_v$-path $p$ of length $l$,
let $i_1, \ldots, i_l$ be the indices of the vertices of this path.
We denote a partition of $p$ into subpaths by the indices of the last vertex in
each subpath i.e. we use $j_1, \ldots, j_k$ to denote the partition 
$((v_{i_1}, \ldots, v_{i_{j_1}}), \ldots, (v_{i_{j_{k-1} + 1}}, \ldots,
v_{i_{j_k}}))$.
Given a partition $(j_1, \ldots, j_k)$ we denote by $V(j_i)$ the vertices of the
$i$th subpath.
A (feasible) \emph{embedding} of a virtual path into the physical graph 
w.r.t partition $P = (j_1, \ldots,j_k)$ is a mapping $\psi : P \to V_p$ such that
for all $1 \leq i \leq k$ it holds that:
\begin{enumerate}
  \item $p_c(\psi(j_i)) \geq \sum_{v \in V(j_i)} p_d(v)$
  \item There is a physical path, $\pi_i$ from $\psi(j_i)$ to $\psi(j_{i+1})$ with a
  bottleneck at least $b_d(v_{j_i}, v_{j_{i+1}})$.
  \item $\bigcup_{i = 1}^k \pi_i$ is a simple path.
\end{enumerate}
The \emph{cost} of a (feasible) embedding, $e_P$ is given by:
$\sum_{i = 1}^k \sum_{v \in V(j_i)} c(v, e_P(j_i))$.

In the \todo{choose a name for the problem} VPN problem the goal is to find a
feasible embedding of a virtual path into the physical DAG that minimizes the
cost.

\begin{figure}[ht]
\centering
\scalebox{1}{
\input{fig-embedding}
}

\end{figure}