### 2017-06-18 src/problem.tex

We are given a virtual directed graph $D_v = (V_v, A_v)$ 
with source and destination vertices $s_v, t_v$ respectively, 
a physical directed graph $D_p = (V_p, A_p)$
with source and destination vertices $s_p, t_p$ respectively,
demand functions $p_d : V_v \to \R$, $b_d : E_v \to \R$,
capacity functions $p_c : V_p \to \R$, $b_c : E_p \to \R$,
and a cost function $c: V_v \times V_p \to \R$.

Given a virtual $s_vt_v$-path $p$ of length $l$,
let $i_1, \ldots, i_l$ be the indices of the vertices of this path.
We denote a partition of $p$ into subpaths by the indices of the last vertex in
each subpath i.e. we use $j_1, \ldots, j_k$ to denote the partition 
$((v_{i_1}, \ldots, v_{i_{j_1}}), \ldots, (v_{i_{j_{k-1} + 1}}, \ldots,
v_{i_{j_k}}))$.
Given a partition $(j_1, \ldots, j_k)$ we denote by $V(j_i)$ the vertices of the
$i$th subpath.
A (feasible) \emph{embedding} of a virtual path into the physical graph 
w.r.t partition $P = (j_1, \ldots,j_k)$ is a mapping $\psi : P \to V_p$ such that
for all $1 \leq i \leq k$ it holds that:
\begin{enumerate}
  \item $p_c(\psi(j_i)) \geq \sum_{v \in V(j_i)} p_d(v)$
  \item There is a physical path, $\pi_i$ from $\psi(j_i)$ to $\psi(j_{i+1})$ with a
  bottleneck at least $b_d(v_{j_i}, v_{j_{i+1}})$.
  \item $\bigcup_{i = 1}^k \pi_i$ is a simple path.
\end{enumerate}
The \emph{cost} of a (feasible) embedding, $e_P$ is given by:
$\sum_{i = 1}^k \sum_{v \in V(j_i)} c(v, e_P(j_i))$.

In the \VPN problem the goal is to find a
feasible embedding of a virtual path into the physical DAG that minimizes the
cost.
Figure~\ref{fig:problem} illustrate an embedding of a virtual request into a
physical network.

\begin{figure}[ht]
\centering
\scalebox{.8}{
\input{fig-embedding}
}
\caption{
An embedding of a graph representation of a virtual request (on the left) into
the physical network (on the right).
The source and destination vertices are dashed, as well as the chosen virtual
path.
The numbers on the vertices represent the processing demand and capacity
respectively, 
the numbers on the edges represent the bandwidth demand and capacity
respectively.
The partition of the virtual path and its embedding are marked by the orange,
dotted arrows.
The cost function is not depicts in this figure.
\label{fig:problem}
}
\end{figure}