### 2017-06-18 src/main.tex
alg

\RequirePackage{amsmath}
\documentclass[runningheads]{llncs}

\input{libs}
\input{def}
\input{tikz}

\input{title}
\input{authors}

\begin{document}
\maketitle

\input{outline}

% \bibliographystyle{plain}
% \bibliography{main}

\end{document}