### 2017-06-18 src/algorithm.tex
alg

In this section we present an FPTAS algorithm to the \VPN{} problem.
We are given an instance of the \VPN{} problem 
\begin{align*}
D_v = (\beta , A_v)		\\
D_p = (v, A_p)		\\
(\sigma , \tau , s, t)	\\
(p_d, b_d, p_c, b_c, c)
\end{align*}
For now, assume all values are integral and polynomial bounded in the size of
the input.
Under this assumption we show that an optimal embedding can be computed
efficiently.
We then argue that, using a standard scaling and rounding technique, we can use
the same computation to achieve a solution with a cost only
$1 + \varepsilon$ times greater that of the optimal solution, for any
$\varepsilon > 0$.
Also, for convenient and w.l.o.g, assume that $\sigma , \tau $ are dummy vertices, 
i.e. $c(\sigma , v) = c(\tau , v) = 0$ for every $v \in v$.

We start by arguing that the minimum processing time required to embed some
$\alpha \gamma $-path into a vertex $v$ can be computed efficiently.
Let $P_{v}(\alpha , \gamma , C)$ denote the minimum processing required to embed
all the virtual vertices, excluding $\alpha $, on a $\alpha \gamma $-path into the physical
vertex $v$ with cost constraint $C$.
$P$ can be computed recursively as follow:
\begin{align*}
&P_{v}(\alpha , \gamma , C) = 
\min_{\alpha \beta  \in A_v} P_{v}(\alpha , \beta , C - c(\gamma , v))
\\
&P(\alpha , \alpha , C) = 0
\\
&P(-, -, <0) = \infty
\end{align*}

We now show how to compute the optimal embedding.
Let $C(\gamma , w)$ denote the
optimal embedding of a $\sigma \gamma $-path into a $sw$-path.
Let $\pi:\beta  \to \N$ be any typological ordering of $\beta $. 
We claim that $O$ can be computed recursively as follow:
\begin{align*}
&C(\gamma , w) = 
\min_{\substack{
\alpha  : pi(\alpha ) \leq pi(\beta ), 
\\
vw \in A_p
}}
C(\alpha , u)
+
\argmin_{C : P_{w}(\beta , \gamma , C) \leq p_c(w)}
\\
&C(\sigma , w) = c(\sigma , w)
\end{align*}
Now, to compute the optimal embedding, one need to compute $C(\tau , t)$. 
We also note that in order to obey to the bandwidth constraints, 
one need to temporally remove all edges $\alpha \beta $ with bandwidth greater than
$b_c(v, w)$ from the virtual graph and only then compute $P_{w}(\beta , \gamma , C)$.
The above formulation is illustrated in Figure~\ref{fig:dp1}.

\begin{figure}[ht]
\centering
\input{fig-dp1}
\caption[]{
\label{fig:dp1}
The optimal embedding of a $\sigma \gamma $-path into a $sw$ path can be efficiently 
computed by breaking the problem into the problem of embedding a $\sigma \alpha $-path
into a $sv$-path (blue, dashed rectangles) 
and the problem of embedding a $\alpha \gamma $-path into $w$ (red, dotted
rectangles) where $b_d(\alpha , \beta ) \leq b_c(v, w)$.
}
\end{figure}