### 2017-06-21 src/graphs.tex

In this section we consider the problem on undirected (physical) graphs.
Once again we seek to embed a virtual $\sigma\tau$-path 
into a simple physical $st$-path.
Note, that we still assume that the virtual graph is a DAG.

We extend our algorithm to work also for this case as well.
Recall that the problem on general graphs cannot be approximate at all,
thus we assume in this section that the there exists an optimal embedding using
at most $k$ \emph{heavy} vertices, thus are vertices with degree greater than two.
The modified algorithm runs in time polynomial in the input length and
exponential in $k$ and achieves the same guarantees as before with high
probability.

The modified algorithm acts as follow:
let $K = {v_1, \ldots, v_k}$ be the set of heavy vertices, and let 
$\pi:[k] \to [k]$ be a random permutation.
Let $v_1, v_l \in K$ be two heavy vertices such that $\pi(v_1) < \pi(v_l)$ and
let $w_1, \ldots, w_l$ any light path between them, i.e. a path that does not
contains any other heavy vertex.
We orient the edges from $u$ to $v$, i.e. $(w_i, w_{i+1})$ for every 
$1 \leq i \leq l - 1$.
One can be verify that this can be done efficiently and that after this process
each edge become an arc and that $G$ is a DAG.

Completing the orientation, we use our previous algorithm to find an optimal
embedding.
We repeat this process, each time with a new independent random permutation, and
keep the best embedding so far.

We claim that with high probability, after $O(k!)$ iterations, the above
algorithm finds the optimal embedding, formally:

\begin{theorem}
With probability at least $1 - e^{-x}$ the algorithm finds the optimal embedding
after $xk!$ iterations.
\end{theorem}

\begin{proof}
Consider an optimal embedding and the order it induces on the heavy vertices,
If the random permutation order the heavy vertices in that order we are
guaranteed to find the optimal embedding.
This happens with probability $1/k!$ and the probability to miss the optimal
embedding after $xk!$ iterations is $1 - (1 - 1/k!)^{xk!} \leq 1 - e^{-x}$.
\end{proof}