### 2017-06-23 src/main.tex

\nonstopmode
%\documentclass[11pt]{article}
\documentclass[runningheads]{llncs}

\usepackage{amsmath,amssymb,mathtools}
%\usepackage{amsthm}
%\usepackage{fullpage}
\usepackage{xspace}
\usepackage{color,soul}
\usepackage{paralist}

% \usepackage[linesnumbered, ruled]{algorithm2e}

\usepackage[final]{microtype}
\usepackage[final]{hyperref}
\usepackage{subcaption}

% DEBUG
%\usepackage{lipsum}
\usepackage{todonotes}
%\usepackage{lineno}
%\linenumbers

% EXTRA
%\usepackage{authblk}

\input{tikz}

\def\VPN{Service Chain Placement in SDN}
\def\R{\mathbb{R}}
\def\N{\mathbb{N}}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% handy commands

\newcommand{\eqdf}{\stackrel{\scriptscriptstyle \triangle}{=}}
\newcommand{\argmin}{\ensuremath{\mbox{argmin}}}
\newcommand{\argmax}{\ensuremath{\mbox{argmax}}}
\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\paren}[1]{\left( #1 \right)}
\newcommand{\bit}{\set{0,1}}
\newcommand{\inv}[1]{\frac{1}{#1}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\ceil}[1]{\left\lceil {#1} \right\rceil}
\newcommand{\floor}[1]{\left\lfloor {#1} \right\rfloor}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\threehalves}{\frac{3}{2}}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}

\newcommand{\eps}{\varepsilon}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\scp}{\textsc{SCP}\xspace}
\newcommand{\scplong}{\textsc{Service Chain Placement}\xspace}

\newcommand{\calE}{\mathcal{E}}
\newcommand{\calG}{\mathcal{G}}
\newcommand{\calV}{\mathcal{V}}

\newcommand{\dror}[1]{\sethlcolor{yellow}\hl{Dror: #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bf Service Chain Placement in SDNs%
\thanks{Research supported in part by Network Programming (Neptune)
  Consortium, Israel.}
}

\titlerunning{Service Chain Placement in SDNs}

%\author[1]{Gilad Kutiel%
%\thanks{E-mail: \texttt{gkutiel@cs.technion.ac.il}}}

%\author[2]{Dror Rawitz%
%\thanks{Partially supported by the Israel Science Foundation
%  (grant no.~497/14).  E-mail: \texttt{dror.rawitz@biu.ac.il}}}

%\affil[1]{Department of Computer Science, Technion, Haifa, Israel}
%\affil[2]{Faculty of Engineering, Bar Ilan University, Ramt-Gan, Israel}

\author{%
Gilad Kutiel
\and
Dror Rawitz%
\thanks{Partially supported by the Israel Science Foundation
  (grant no.~497/14).  }
}

\institute{%
Department of Computer Science, Technion, Haifa 32000, Israel. \\
\email{gkutiel@cs.technion.ac.il}
\and
Faculty of Engineering, Bar Ilan University, Ramt-Gan 52900, Israel.\\
\email{dror.rawitz@biu.ac.il}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

% \begin{abstract}
% We study the allocation problem of a \emph{service chain} in a
% software-defined network that supports network function
% virtualization.
% %
% Given a network that contains servers with limited processing power
% and of links with limited bandwidth, a service chain is a sequence of
% virtual network functions (VNFs) that service a certain flow request
% in the network.
% %
% The allocation of a service chain consists of routing and VNF
% placement.  That is, each VNF from the seqeunce is placed in a server
% along a path, and it is feasible if each server can handle the VNFs
% that are assigned to it, and if each link on the path can carry the
% flow that is assigned to it.
% %
% A request for service is composed of a source and a destination in the
% network, an upper bound on the total latency, and a specification of
% all service chains that are considered valid for this request.
% %
% Each pair of server and VNF are associated with a cost for placing the
% network function in the server, and given a request, the goal is to
% find a valid service chain of minimum total cost or to identify that
% a valid service chain does not exist.
% 
% We show that even the feasibility problem is NP-hard in general
% di-graphs.  Hence we first focus on directed acyclic graphs (DAGs).
% We show that the problem is still NP-hard in DAGs and present an
% FPTAS.
% %
% Based on out PTAS, we also provide a randomized algorithm for
% instances in which the service chain passes through a bounded number
% of servers whose degree is larger than two.
% 
% ?????
% 
% Fault tolerant placement: two service chains?
% 
% FPT algorithm where the number of servers whose degree is larger than
% two is a parameter?
% \end{abstract}
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \section{Introduction}
% 
% 
% Computer communication networks are in constant need of expansion to
% cope with the ever growing traffic.  As networks grow, management and
% maintenance become more and more complicated.
% %
% Current developments aim to improve the utilization of network
% resources include the detachment of network applications from network
% infrastructure and the transition from network planning to network
% programming.
% 
% One aspect of network programming is to manage resources from a
% central point of view, namely to make decisions based on availability,
% network status, required quality of service, and the identity of the
% client.  Hence, a central issue is a central agent that is able to
% received reports from network components and client requests, and as a
% result can alter the allocation of resources in the networks.  This
% approach is called Software Defined Networking (SDN), where there is a
% separation between routing and management (control plane) and the
% underlying routers and switches that forward traffic (data plane) (see
% Kreutz et al.~\cite{KRVRAU15}).
% 
% A complimentary approach is Network Function Virtualization
% (NFV)~\cite{NFV12}.  Instead of relying on special purpose machines,
% network applications become virtual network functions (VNF) that are
% executed on generic machines and can be placed in various locations in
% the network.  Virtualization increases the flexibility of resource
% allocation and thus the utilization of the network resources.
% %
% Internet Service Providers (ISPs) that provide services to clients
% benefit from NFV, since it helps to better utilize their physical
% network.  In addition, when network services are virtualized, an ISP
% may support \emph{service chaining}~\cite{ServiceChaining15}, namely a
% compound service that consists of a sequence of VNFs.  Furthermore,
% one may offer a compound service a service that may be obtained using
% one of several sequences of VNFs.
% 
% In this paper we study the allocation problem of a compund request for
% a service chain in a software-defined network that supports NFV.
% Given a (physical) network that contains servers with limited
% processing power and of links with limited bandwidth, a service chain
% is a sequence of virtual network functions (VNFs) that service a
% certain flow request in the network.
% %
% The allocation of a service chain consists of routing and VNF
% placement.  That is, each VNF from the seqeunce is placed in a server
% along a path, and it is feasible if each server can handle the VNFs
% that are assigned to it, and if each link on the path can carry the
% flow that is assigned to it.
% %
% A request for service is composed of a source and a destination in the
% network, an upper bound on the total latency, and a specification of
% all service chains that are considered valid for this request.
% %
% Each pair of server and VNF are associated with a cost for placing the
% network function in the server.  This cost measures compatibility of a
% VNF to a server (infinite costs means ``incompatible'').  Given a
% request, the goal is to find a valid service chain of minimum total
% cost or to identify that a valid service chain does not exist.
% 
% 
% \dror{Explain one shot approach}
% 
% \paragraph*{\bf Related work.}
% %
% 
% Cohen et al.~\cite{CLNR15} considered VNF placement.  In their model
% the input is an undirected graph with a metric distance function on
% the edges.  clients are located in various nodes, and each client is
% interested in a subset of VNFs.  For each node $v$ and VNF $\alpha$,
% there is a setup cost associated with placing a copy of $\alpha$ at
% $v$ (multiple copies are allowed), and there is a load that is induced
% on $v$ for placing a copy of $\alpha$.  Furthermore, each node has a
% limited capacity.  Also, each copy of a VNF can handle a limited
% amount of clients A solution is the assignment of each client to a
% subset of nodes, each corresponding to one of its required VNFs.  The
% cost of a solution is the total setup costs plus the sum of distances
% between the clients and the node from which they get service.
% %
% Cohen et al.~\cite{CLNR15} gave bi-criteria approximation algorithms
% for various versions of the problem, namely algorithms that compute
% constant factor approximations that violate the load constraints by a
% constant factor.
% %
% It is important to note that in this problem does not consider
% routing, and also it assumes unordered VNF subsets.
% 
% Even, Medina, and Patt-Shamir~\cite{EMP16} studied online path
% computation and VNF placement.  They considered service chain requests
% that arrive in an online manner.  Each request is a a flow with a
% high-level specification of routing and VNF requirements.  Each VNF
% can be performed by a specified subset of servers in the system.  Upon
% arrival of a request, the algorithm either rejects the request or
% accepts it and with a specific routing and VNF assignment, under given
% server capacity constraints.  Each request has a benefit that is
% gained if it is accepted, and the goal is to maximize the total
% benefit.
% %
% Even, Medina, and Patt-Shamir~\cite{EMP16} proposed an algorithm that
% copes with requests with unknown duration without preemption by using
% a third response, which is refer to as “stand by”, whose competitive
% ratio is $O(\log (knb_{\max})$, where $n$ is the number of nodes,
% $b_{\max}$ is an upper bound on a request benefit per time unit, and
% $k$ is the maximum number of VNFs of any service chain.  This
% performance guarantee holds for the case where the processing of any
% processing request in any possible service chain is never more than
% $O(1/(k \log (nk)))$ fraction of the capacity of any network
% component.
% 
% Even, Rost, and Schmid~\cite{ERS16}
% 
% \dror{finish}
% 
% \paragraph*{\bf Our results.}
% %
% We show that even the feasibility problem is NP-hard in general
% di-graphs.  Hence we first focus on directed acyclic graphs (DAGs).
% We show that the problem is still NP-hard in DAGs and present an
% FPTAS.
% %
% Based on out PTAS, we also provide a randomized algorithm for
% instances in which the service chain passes through a bounded number
% of servers whose degree is larger than two.
% 
% ?????
% 
% \bigskip
% 
% 
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \section{Preliminaries}
% 
% 
% \paragraph*{\bf Model.}
% %
% The \scplong (\scp) input is composed of three components, a physical
% network, a virtual specification, and placement costs:
% \begin{description}
% \item[Physical network:]
% %
%   The physical network is a digraph $G = (V,E)$.  Each node $v \in V$
%   has a non-negative processing capacity $p(v)$, and each directed
%   edge $e \in A$ has a non-negative bandwidth cap $b(e)$.
% 
% \item[Virtual specification:]
% %
%   The description of a request for a service chain consists of a
%   physical source $s \in V$, a physical destination $t \in V$, and a
%   directed acyclic graph (DAG) $\calG = (\calV,\calE)$.
% %
%   Without loss of generality, we assume that $p(s) = p(t) = 0$.  
% %
%   The DAG $\calG$ has a source node $\sigma \in \calV$ and a
%   destination $\tau \in \calV$ Each node $\alpha \in \calV$ represent
%   a VNF that has a processing demand $p(\alpha)$.  Without loss of
%   generality, we assume that $p(\sigma) = p(\tau) = 0$.  Each edge
%   $\eps \in \calE$ has a bandwidth demand $b(\eps)$.
%   
% \item[Placement costs:]
% %
%   There is a non-negative cost $c(\alpha,v)$ for placing the VNF
%   $\alpha$ in $v$.  We assume that $c(\sigma,s) = 0$ and $c(\sigma,v)
%   = \infty$, for every $v \neq s$.  Similarly, we assume that
%   $c(\tau,t) = 0$ and $c(\tau,v) = \infty$, for every $v \neq t$.
% \end{description}
% 
% A solution consists of the following:
% \begin{description}
% \item[Virtual path:] A path from $\sigma$ to $\tau$ in $\calG$, namely
%   a sequence of vertices $\alpha_0,\ldots,\alpha_\ell$, where
%   $\alpha_0 = \sigma$, $\alpha_\ell = \tau$, and
%   $(\alpha_j,\alpha_{j+1}) \in \calE$, for every $j \in
%   \set{0,\ldots,\ell-1}$.
% 
% \item[Physical path:] A simple path from $s$ to $t$ in $G$, namely a
%   sequence of node $v_0,\ldots,v_k$, where $v_0 = s$, $v_k = t$, and
%   $(v_i,v_{i+1}) \in E$, for every $i \in \set{0,\ldots,k-1}$.
%   
% \item[Placement:] A function $f$ that maps the virtual path to the
%   physical path.  Formally, a placement is a function $f:
%   \set{\alpha_0,\ldots,\alpha_\ell} \to \set{v_0,\ldots,v_k}$
%   such that
% \begin{enumerate}
% \item If $f(\alpha_j) = v_i$ and $f(\alpha_{j+1}) = v_{i'}$, then $i
%   \leq i'$.
% \item $b(\alpha_j,\alpha_{j+1}) \leq b(v_i,v_{i+1})$, for every
%   $(v_i,v_{i+1}) \in \bar{f}(\alpha_j,\alpha_{j+1})$, where
%   $\bar{f}(\alpha_j,\alpha_{j+1})$ to be the set of physical edges
%   that correspond to it, i.e.,
% \[
% \bar{f}(\alpha_j,\alpha_{j+1}) \eqdf \set{(v_i,v_{i+1}) : i' \leq i < i''}
% ~,
% \]
%   where $f(\alpha_j) = v_{i'}$ and $f(\alpha_{j+1}) = v_{i''}$.
% 
% \item $\sum_{\alpha \in f^{-1}(v)} p(\alpha) \leq p(v)$, where
%   $f^{-1}(v) \eqdf \set{\alpha : f(\alpha) = v}$.
%  
% \end{enumerate}
% \end{description}
% 
% An example of an \scp instance and a solution is given in
% Figure~\ref{fig:solution}.
% 
% \begin{figure}
% \centering
% \scalebox{1}{
% \input{fig-embedding}
% }
% \caption{???}
% \label{fig:solution}
% \end{figure}
% 
% The cost of a placement $f$ is:
% \[
% \sum_j c(\alpha_j,f(\alpha_j))
% ~.
% \]
% In \scp the goal is to find a feasible placement of a virtual path
% into the physical DAG that minimizes the cost.
% 
% %%%%%
% 
% We also consider an extended version of \scp in which each physical
% link causes a delay and there is an upper bound $T$ on the total
% delay.  More formally, each pair of a virtual edge $\eps$ and physical
% edge $e$ is associated with a latency $t(\eps,e)$.
% %
% Given a placement $f: \set{\alpha_0,\ldots,\alpha_\ell} \to
% \set{v_0,\ldots,v_k}$, the total latency of the solution is given by
% \[
% T(f) = \sum_j \sum_{e \in \bar{f}(\eps)} t(\eps,e)
% ~.
% \]
% In this case there is an additional constraint that $T(f) \leq T$.
% 
% 
% %%%%%
% 
% \paragraph*{\bf Definitions and notation.}
% %
% Define $n_p \eqdf \abs{V}$ and $m_p \eqdf \abs{E}$.  Similarly, define
% $n_v \eqdf \abs{\calV}$ and $m_v \eqdf \abs{\calE}$.
% 
% Given a virtual DAG $\calG$, we assume the existence of a topological
% sorting of the vertices and write $\alpha \prec_v \beta$ if $\alpha$
% precedes $\beta$ in this ordering.  If the physical network is a DAG,
% then we make a similar assumption and write $v \prec_p u$ if $v$
% precedes $u$ in this ordering.
% 
% 
% 
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \section{Hardness}
% 
% In this section we present two hardness results.  First, we show that
% even the feasibility question of \scp is NP-hard, and therefore no
% approximation algorithm exists for \scp.  In addition, we show that
% \scp is NP-hard even if $\abs{V \setminus \set{s,t}} = 1$.  We also
% show that \scp is NP-hard even if both the physical network and the
% virtual DAG are simple paths.
% 
% We show that finding a feasible solution is NP-hard.
% 
% \begin{theorem}
% Feasibility of \scp is NP-hard
% \end{theorem}
% \begin{proof}
% We use a reduction from \textsc{Hamiltonian Path} that is known to be
% NP-hard~\cite{GarJoh79}.
% %
% Given an instance $H$ of \textsc{Hamiltonian Path} we construct an
% instance of \scp.  For the physical network we have that $G$, where
% $V(G) = V(H) \cup \set{s,t}$, and $E(G) = E(H) \cup \set{(s,v),(v,t) :
%   v \in V(H)}$.  In addition, $p(v) = 1$, for every $v$, and $b(e) =
% 1$, for every $e$.  The virtual DAG $\calG$ is a path containing $n+2$
% virtual functions, $\sigma = \alpha_0, \alpha_1, \ldots, \alpha_n,
% \alpha_{n+1} = \tau$, where $p(\alpha_i) = 1$, for every $i \in
% \set{1,\ldots,n}$.  Also, $b(\alpha_i,\alpha_{i+1}) = 1$, for every
% $i$.
% 
% The construction can clearly be computed in polynomial time.
% %
% An Hamiltonian Path in $G$, induces a \scp solution that follows the
% path, i.e., $\alpha_i$ is placed in the $i$th vertex along the path.
% %
% One the other hand, since all demands and capacities are $1$, no two
% VNFs can share a physical node.  Hence a \scp solution induces an
% Hamiltonian path.
% %
% \qed
% \end{proof}
% 
% Next, we show that \scp is NP-hard even if the physical network is a
% DAG.
% 
% \begin{theorem}
% \scp is NP-hard, even if $\abs{V \setminus \set{s,t}} = 1$.
% \end{theorem}
% \begin{proof}
% The prove the theorem using a reduction from \textsc{Partition}.
% Given a \textsc{Partition} instance $\set{a_1, \ldots, a_n}$ we
% construct an \scp instance as follows.
% %
% The physical network $G$ contains three nodes: $s$, $v$, and $t$, and
% there are two edges $(s,v)$ and $(v,t)$.  The capacity of $v$ is $p(v)
% = \half \sum_i a_i$.  Edge capacities are zero.
% %
% The virtual DAG is composed of $2n+2$ vertices, namely
% \[
% \calV
% = \set{\sigma,\tau}
%   \cup \set{\alpha_1, \ldots, \alpha_n}
%   \cup \set{\beta_1, \ldots, \beta_n}
% ~.
% \]
% Also,
% \[
% \calE
% = \bigcup_i \set{ \set{\alpha_i,\beta_i} \times \set{\alpha_{i+1},\beta_{i+1}} } 
%   \cup
%   \set{ (\sigma,\alpha_1), (\sigma,\beta_1), (\alpha_n,t), (\beta_n,t) }
% ~.
% \]
% The DAG is shown in Figure~\ref{fig:dag}.
% %
% The demands are $p(\alpha_i) = a_i$ and $p(\beta_i) = 0$, for every
% $i$.  Also, $b(\eps) = 0$, for every $\eps \in \calE$.  The costs are:
% $c(\alpha_i,v) = 0$ and $c(\beta_i,v) = a_i$, for every $i$.
% 
% The \scp instance can be computed in polynomial time.
% %
% Also, it is not hard to verify that $a \in \textsc{Partition}$ if and
% only if there exists a solution whose cost is $\half \sum_i a_i$.
% %
% \qed
% \end{proof}
% 
% \begin{figure}
% \centering
% \input{fig-reduction}
% \caption{???}
% \label{fig:dag}
% \end{figure}
% 
% \dror{Figure 2: source and destination should be fixed}
% 
% \begin{theorem}
% \scp is NP-hard, even if both the physical network and the virtual DAG
% are paths.
% \end{theorem}
% \begin{proof}
% 
% \dror{reduction from partition using latency}
%   
% %
% \qed
% \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Algorithm}
\input{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Latency}
\input{latency}

\section{Undirected Graphs}
\input{graphs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{chain}

\end{document}