### 2017-06-23 src/fig-reduction2.tex

\begin{tikzpicture}[every node/.style={default node}]
\node[draw=none] at(3, 1) {Virtual Path};

\node(alpha1) at(0,0) {$\alpha_1$};
\node(alpha2) at(2,0) {$\alpha_2$};
\node[draw=none] at(6,0) {$\cdots$};
\node(alphan) at(10,0) {$\alpha_n$};
\node(tau) at(12,0) {$\tau$};

\draw (alpha1) -- (alpha2);
\draw (alpha2) -- +(2, 0);
\draw ($(alphan) + (-2,0)$) -- (alphan);
\draw (alphan) -- (tau);

\node(v1-) at(0, -2) {$v_1^-$};
\node(v1+) at(0, -4) {$v_1^+$};
\node(v2-) at(2, -2) {$v_2^-$};
\node(v2+) at(2, -4) {$v_2^+$};

\node[draw=none] at(6,-2) {$\cdots$};
\node[draw=none] at(6,-4) {$\cdots$};

\node(vn-) at(10, -2) {$v_n^-$};
\node(vn+) at(10, -4) {$v_n^+$};
\node(t) at(12, -2) {$t$};

\draw (v1-) -- (v1+) node[label above] {$a_1$};
\draw (v1+) -- (v2-) node[label below] {$0$};
\draw (v2-) -- (v2+) node[label above] {$a_2$};
\draw (v2+) -- +(2, 2)node[label below] {$0$};

\draw ($(vn-) + (-2,-2)$) -- (vn-)node[label below] {$0$};;
\draw (vn-) -- (vn+)node[label above] {$a_n$};
\draw (vn+) -- (t)node[label below] {$0$};;

\begin{scope}[every node/.style={label above, orange}]
\draw[group, ->] (alpha1) to[bend left] node {$0$} (v1-);
\draw[group, ->] (alpha1) to[bend right] node {$a_1$} (v1+);
\draw[group, ->] (alpha2) to[bend left] node {$0$} (v2-);
\draw[group, ->] (alpha2) to[bend right] node {$a_2$} (v2+);
\draw[group, ->] (alphan) to[bend left] node {$0$} (vn-);
\draw[group, ->] (alphan) to[bend right] node {$a_n$} (vn+);

\draw[group, ->] (tau) to[bend left] node {$0$} (t);
\end{scope}

\end{tikzpicture}